---
title: "Select Top *n* Target Locus Candidates from lDS004"
author: "Katya Williams"
date: "`r Sys.Date()`"
output: html_document
---

# 1. Setup

## 1.1 Import necessary libraries

```{r setup}
# Load any necessary libraries
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(seqinr)
```

## 1.2 Read in lDS004 datasets

```{r read-lDS004}
# read in lDS004 data (from Ann Cirincione)
lds004 <- read_csv("lDS004_only_d7.csv")          # contains info on the edit
lds004_avg <- read_csv('lDS004_avg_outcomes.csv') # contains editing percentages

# select relevant columns
selected <- lds004 %>% select(pegRNA_ID, cell_line, Gene, Guide, PBS, RT_template, Barcode,
                              Wide_target_sequence, Reference_nt, Edited_nt, Protospacer,
                              read_pct_d7, PBS_length, RT_length, GC_percent_PBS)
selected_avg <- lds004_avg %>% select(pegID, CellLine, Day7.Avg.Freq.TotalEdited,
                                      Day7.Avg.Freq.EditOnly, Day7.Avg.Freq.TotalErrors,
                                      Day7.Avg.TotalReads)
```

## 1.3 Read in any auxiliary datasets

```{r read-aux}
# read in list of essential genes from depmap and set up data for use
## 'common_essentials.csv' from depmap.org
essential <- read_csv('common_essentials.csv')
essential <- essential %>% separate(gene, into = c("gene", "num"), sep = "\\s")
```

# 2. MMR filtering

## 2.1 Rank by sensitivity to MMR loss

```{r MMR-rank}
# set up data to create separate cell line columns for editing percentages
## arrange so that biggest group is gene, next biggest group is edited nt, smallest group is cell line
sort <- selected %>% arrange(cell_line) %>% arrange(Edited_nt) %>% arrange(Gene)
## group 004_avg by pegID
sort_avg <- selected_avg %>% arrange(pegID)

# generate editing difference between WT and MMR KO
## create new sets, one for each cell line
mlh1 <- sort %>% filter(cell_line == 'MLH1_KO')
pemax <- sort %>% filter(cell_line == 'PEmax')
mlh1_avg <- sort_avg %>% filter(CellLine == 'MLH1_KO')
pemax_avg <- sort_avg %>% filter(CellLine == 'PEMax')

## join back together by pegRNA ID
wide <- right_join(mlh1, pemax, by=c('pegRNA_ID' = 'pegRNA_ID'))
wide_avg <- right_join(mlh1_avg, pemax_avg, by=c('pegID' = 'pegID'))

## give columns more meaningful names
wide <- wide %>% rename(Gene = Gene.x, Guide = Guide.x, PBS = PBS.x, 
                        RT_template = RT_template.x, Barcode = Barcode.x,
                        Wide_target_sequence = Wide_target_sequence.x, 
                        Reference_nt = Reference_nt.x, Edited_nt = Edited_nt.x, 
                        PBS_length = PBS_length.x, RT_length = RT_length.x, 
                        GC_percent_PBS = GC_percent_PBS.x, Protospacer = Protospacer.x)
wide_avg <- wide_avg %>% rename(MLH1_Avg_Freq_TotalEdited_d7 = Day7.Avg.Freq.TotalEdited.x, 
                                PEmax_Avg_Freq_TotalEdited_d7 = Day7.Avg.Freq.TotalEdited.y,
                                MLH1_Avg_Freq_EditOnly_d7 = Day7.Avg.Freq.EditOnly.x, 
                                PEmax_Avg_Freq_EditOnly_d7 = Day7.Avg.Freq.EditOnly.y, 
                                MLH1_Avg_Freq_TotalErrors = Day7.Avg.Freq.TotalErrors.x, 
                                PEmax_Avg_Freq_TotalErrors = Day7.Avg.Freq.TotalErrors.y,
                                MLH1_Avg_TotalReads = Day7.Avg.TotalReads.x, 
                                PEmax_Avg_TotalReads = Day7.Avg.TotalReads.y)

# keep only the important columns in wide, wide_avg
wide <- wide %>% select(pegRNA_ID, Gene, Guide, PBS, RT_template, Barcode, Wide_target_sequence,
                        Reference_nt, Edited_nt, PBS_length, RT_length,
                        GC_percent_PBS, Protospacer)
wide_avg <- wide_avg %>% select(pegID,
                                MLH1_Avg_Freq_TotalEdited_d7, PEmax_Avg_Freq_TotalEdited_d7,
                                MLH1_Avg_Freq_EditOnly_d7, PEmax_Avg_Freq_EditOnly_d7, 
                                MLH1_Avg_Freq_TotalErrors, PEmax_Avg_Freq_TotalErrors,
                                MLH1_Avg_TotalReads, PEmax_Avg_TotalReads)

## calculate the difference and proportion in editing percentage by cell line
wide_avg <- wide_avg %>% mutate(mean_diff_avg_freq = MLH1_Avg_Freq_EditOnly_d7 - PEmax_Avg_Freq_EditOnly_d7)

# join data together by pegRNA ID
diff <- full_join(wide, wide_avg, by=c('pegRNA_ID' = 'pegID'))
```

## 2.2 Remove negative controls and essential genes

Since the pegRNA will edit the gene in its original context, ensure the pegRNA will not modify genes with essential functions -- this will give us a better estimate of editing efficacy since cells will be less likely to die as a direct result of editing.

```{r remove-essential}
# remove unhelpful data
## filter out any non-targeting rows
diff <- diff[!grepl('non-targeting', diff$Gene),]

## remove any essential genes
### uses essential genes dataset read in in 1.3
diff <- diff %>% filter(!Gene %in% essential$gene)
```

## 2.3 Rank MMR sensitivity by approximating the maximum difference between MLH1 KO and PEmax efficiencies

```{r model}
# create a model to use to approximate max difference between MLH1 KO and PEmax
## fit the model
model <- loess(diff$MLH1_Avg_Freq_EditOnly_d7 ~ diff$PEmax_Avg_Freq_EditOnly_d7, na.action=na.exclude)

## store the predicted MLH1 KO alue for each PEmax value per row
prediction <- predict(model)

## find the residuals and store, then arrange by smallest residuals first
diff <- diff %>% mutate(Residuals = diff$MLH1_Avg_Freq_EditOnly_d7 - prediction) %>% 
    arrange(Residuals)
```

## 2.4 Select number of edits

*Modify edit tuning parameters min_MLH1, max_PEmax to change edit selection number*

```{r tuning-parameters}
# tuning parameters for selection
## number of edit candidates to work with: change to optimize library size (Tier 1: end goal of max 100)
## minimum threshold for MLH1 KO efficiency
## maximum threshold for PEmax WT efficiency
min_MLH1 <- 0.6863 # adjusting up decreases library size
max_PEmax <- 0.094 # adjusting down decreases library size
```

```{r top-from-diff-avg}
# remove any duplicate entries
## create a row identifier of gene + target locus (creates a duplicate gene column, which we then need to remove)
diff <- diff %>% separate(pegRNA_ID, into = c("gene", "Gene_Locus"), sep = "_", extra='drop') %>%
    select(-gene) %>% mutate(Target_Locus = paste(Gene, Gene_Locus, sep="_"))
nrow(diff)
## remove any duplicate entries under Target_Locus, arranged so that the entry with the highest mean_diff_avg_freq remains
diff <- diff %>% arrange(-mean_diff_avg_freq)
diff <- diff[!duplicated(diff$Target_Locus), ]
nrow(diff)

# select n edit candidates by first filtering for edits within the thresholds, then ordering by smallest residual
top <- diff %>% filter(MLH1_Avg_Freq_EditOnly_d7 >= min_MLH1) %>% 
    filter(PEmax_Avg_Freq_EditOnly_d7 <= max_PEmax) %>% 
    arrange(Residuals) 
```

# 3. Export files

## 3.1 Export target loci data

```{r export}
# export files
write.csv(top, 'MMR-responsive_any_target_candidates.csv', row.names=TRUE)
write.csv(diff, 'lDS004_joined_any_mean.csv')
```